---
title: "Week 10: Title here"
author: "Juanjo Medina and Reka Solymosi"
output: html_document
---

##10.1 Introduction



Before we do any of this, we need to load the libraries we will use today:

```{r}
library(sf)
library(tmap)
library(sp)
library(spdep)
```


Then we will bring back the data from last week:

```{r}
##R in Windows have some problems with https addresses, that's why we need to do this first:
urlfile<-'https://s3.amazonaws.com/geoda/data/ncovr.zip'
download.file(urlfile, 'ncovr.zip')
#Let's unzip and create a new directory (ncovr) in our working directory to place the files
unzip('ncovr.zip', exdir = 'ncovr')

```

Last week we did not treated the data as spatial and, consequently, relied on the csv file. But notice that in the unzip ncovr file there is also a shapefile that we can load as a spatial object into R:

```{r}
shp_name <- "ncovr/ncovr/NAT.shp"
ncovr_sf <- st_read(shp_name)
```

#10.2 Interpreting SPatial Lag COefficiences

##Model from HW

PUT RECAP HERE

##impacts()

take from rpub [here](https://rpubs.com/corey_sparks/109650)

##10.3 Cartograms

##10.4 bins / hexbins

##10.5 MAUP

